---
title: "Red Wine Quality"
author: "Arpit Kanodia"
date: "August 14, 2015"
output: word_document
---
```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
library(GGally)
library(scales)
library(memisc)
library(reshape)
library(gridExtra)
```


Importing the CSV file

```{r echo=TRUE, Load_the_Data}
# Load the Data
dataframe <- read.csv('wineQualityReds.csv')
```

## About the Project

In this project, I will explore the quality of wine wrt different properties and chemicals. I will produce the different plots, tables and numeric stats value for finally defining the results.

The data wineQualityReds.csv is available on this link [wineQualityReds.csv](https://www.google.com/url?q=https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityReds.csv&sa=D&usg=AFQjCNFDGQM6iS1viDgkVnXfjlJUymYerA)

And the details about the data is available here [Text file about Details of Data](https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityInfo.txt)


## Basic Details about the Dataset 

Lets examine the structure and schema of dataset

```{r echo=TRUE}
str(dataframe)
summary(dataframe)
```

As our main goal of this project is quality, it will be intresting to see statistics of only quality.

```{r echo=TRUE}
str(dataframe$quality)
summary(dataframe$quality)
```

Some initial basic details of the dataset are observed as following:

- There are 1599 obs. with 13 variables.
- From the literature, the value of `quality` is in range of 0-10, and was rated by atleast 3 whine experts. In the dataset value range from 3-8, with median of 6 and mean at 5.636.
- `X` appears to be a identifier for each data.

```{r echo=FALSE, results='hide', basic_observations}
#lets clean up and explore the data
dataframe$X = factor(dataframe$X)

#quality
summary(dataframe$quality)
table(dataframe$quality)

#check the assertions
dataframe$quality <- factor(dataframe$quality, ordered = T)
str(dataframe$quality)
```

## Plots of Single Variables 



Just to explore the data visully, I'll draw all the graphs for 12 variables quickly. The motive behind to see the distribution. Histograms are required to analyze initially about the distribution of the different variables.

```{r echo=FALSE, message=FALSE, warning=FALSE,single_varable_plot}

fixed_acidity <- qplot(dataframe$fixed.acidity, 
                       color = I('black'), fill = I('#F79420'))
volatile_acidity <- qplot(dataframe$volatile.acidity, 
                          color = I('black'), fill = I('#F79420'))
citric_acid <- qplot(dataframe$citric.acid, 
                     color = I('black'), fill = I('#F79420'))
residual_sugar <- qplot(dataframe$residual.sugar,
                        color = I('black'), fill = I('#F79420'))
chlorides <- qplot(dataframe$chlorides, 
                   color = I('black'), fill = I('#F79420'))
free_sulfur_dioxide <- qplot(dataframe$free.sulfur.dioxide, 
                   color = I('black'), fill = I('#F79420'))
total_sulfur_dioxide <- qplot(dataframe$total.sulfur.dioxide, 
                   color = I('black'), fill = I('#F79420'))
density <- qplot(dataframe$density, 
                   color = I('black'), fill = I('#F79420'))
ph <-  qplot(dataframe$pH, 
                   color = I('black'), fill = I('#F79420'))
sulphates <- qplot(dataframe$sulphates, 
                   color = I('black'), fill = I('#F79420'))
alcohal <- qplot(dataframe$alcohol, 
                 color = I('black'), fill = I('#F79420'))
quality <- qplot(dataframe$quality, 
                 color = I('black'), fill = I('#F79420'))

grid.arrange(volatile_acidity,
             citric_acid,density,
             fixed_acidity,
             free_sulfur_dioxide,
             ph,
             quality,
             residual_sugar,
             sulphates,
             total_sulfur_dioxide,
             volatile_acidity,
             chlorides,ncol=4)
```

## Analysis of Single Variable 

## Quality of the Wine

The quality of wine is in discrete value range from 0-10 as mentioned in description text of dataset. but with graph of quality one can definetely see the quality is in between 3-8, where as most of the data lie between 5-7, with very small data at 3,4 and at 8.
```{r echo=FALSE, message=FALSE, wine_quality_plot}
qplot(dataframe$quality, color = I('black'), fill = I('#F79420'))
```
      
For rating of wine, I thought to divide the quality into 'bad','average', 'good' and 'very good'. As the data range only from 3-8, dividing the data into 4 parameters, whereas 'very good' is non existent as there are no data for 9 and 10, is nonsensical. I finally decided to divide the data into only 3 parameters 'bad', 'average' and 'good'.

```{r echo=TRUE, message=FALSE,rating}
#Dividing data according to Quality
dataframe$rating <- ifelse(dataframe$quality < 5, 'bad', 
                           ifelse(dataframe$quality < 7, 'average', 'good'))
#Adding levels
dataframe$rating <- ordered(dataframe$rating,
                     levels = c('bad', 'average', 'good'))
summary(dataframe$rating)
qplot(dataframe$rating, color = I('black'),fill = I('#F79420'))
```

##Analysis of Distribution

- It is obviously appearing in graphs that maybe pH and density are normally distributed.
-`Sulphur dioxide`,`sulphates`,`alcohol`, `fixed and volatile acidity` looks like to be long tailed.
- Residual sugar and chlorides have a extreme outliers.
- Citric acid seems to have large number of zero values.

#### Log based Plots

To furthur analyze the distribution, log10 based transformation is required.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = dataframe,
       aes(x = fixed.acidity)) +
  geom_histogram(color = I('black'), fill = I('#F79420')) +
  scale_x_log10()

ggplot(data = dataframe,
       aes(x = volatile.acidity)) +
  geom_histogram(color = I('black'), fill = I('#F79420')) +
  scale_x_log10()

ggplot(data = dataframe,
       aes(x = citric.acid)) +
  geom_histogram(color = I('black'), fill = I('#F79420')) +
  scale_x_log10()
```

When plots are made on the base of log 10, `fixed.acidity` and `volatile.acidity` appeared to be normally distributed, as the pH is normally distributed which is measure of acidity in log scale. But, `citric.acid` not appeared to be normally distributed. On furthur observations.

```{r}
length(subset(dataframe, citric.acid == 0)$citric.acid)
```
This is clear that 132 observations are actually zero. 


```{r echo=FALSE, message=FALSE}
p1 <- ggplot(data = dataframe, aes(x = residual.sugar)) +
  geom_histogram(color = I('black'), fill = I('#F79420')) + 
  scale_x_continuous(lim = c(0,quantile(dataframe$residual.sugar, 0.95))) +
  xlab('95th percentile trucated data of Residual Sugar')

p2 <- p1 + scale_x_log10() + 
  xlab('Log10 of 95th percentile trucated data of Residual Sugar')

grid.arrange(p1,p2,ncol=1)

p3 <- ggplot(data = dataframe, aes(x = chlorides)) +
  geom_histogram(color = I('black'), fill = I('#F79420')) + 
  scale_x_continuous(lim = c(0, quantile(dataframe$chlorides, 0.95))) +
  xlab('95th percentile trucated data of Chlorides')

p4 <- p3 + scale_x_log10() + 
  xlab('Log10 of 95th percentile trucated data of Chlorides')

grid.arrange(p3,p4,ncol=1)

p5 <- ggplot(data = dataframe, aes(x = sulphates)) +
  geom_histogram(color = I('black'), fill = I('#F79420'))  +    scale_x_continuous(lim = c(0, quantile(dataframe$sulphates, 0.95))) +
  xlab('95th percentile trucated data of Sulphates')

p6 <- p5 + scale_x_log10() + 
  xlab('Log10 of 95th percentile trucated data of Sulphates')
grid.arrange(p5,p6,ncol=1)

```

##Checking the Overall picture of Dataset 

```{r echo=FALSE, warning=FALSE}
wine_tmp = dataframe[,2:13]
colnames(wine_tmp) = c("Fix.Ac", 
                       "Vol.Ac", 
                       "Citric", 
                       "Sugar", 
                       "Cl", 
                          
                       "Free.SO2", 
                       "SO2", 
                       "Density", 
                       "pH", "SO4", 
                       "Alc",
                       "Quality")
suppressMessages(ggpairs(wine_tmp,  params=list(size=2,binwidth=5)) + 
  theme(axis.text = element_blank()))

rm(wine_tmp)



```


##Some Short Questions

#### Did you create any new variables from existing variables in the dataset?

Yes, `rating` based on quality is created that is bad,average and good wine. In last plot I also created `AlcohalRating` according to % of Alcohol in Wine. 

Furthur I created a new variable for acidity, it seems like `fixed.acidity` and `volatile.acidity` caused because of different acids in the Wine. I created a new `Total.acidity` containing the sum of `fixed.acidity`,`volatile.acidity` and `citric.acidity`.

```{r echo=FALSE}
dataframe$Total.acidity <- dataframe$fixed.acidity + 
  dataframe$volatile.acidity + 
  dataframe$citric.acid

plot <- qplot(dataframe$Total.acidity, 
      color = I('black'), fill = I('#F79420'))
suppressMessages(print(plot
                       
                       ))
```

I searched on internet and I found fixed acidity caused because of tartaric, malic, citric, and succinic acids in wine. And volatile acidity causes because of acetic acid.

Source:

[Fixed Acidity](http://waterhouse.ucdavis.edu/whats-in-wine/fixed-acidity)
[Volatile Acidity](http://waterhouse.ucdavis.edu/whats-in-wine/volatile-acidity)

#### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

I investigated the distribution in above sections. For outliers I decided to plot the boxplots. In Single Variable plot, I didnt tidy or adjust the data, but plotted some on log scale.

####Boxplots of Single Variable

For defining the median across `quality` and `rating` it is required to plot the Boxplot.
```{r,echo=FALSE}
#Creating Boxplot Function for Single Variable
boxplot_function <- function(column, ylab) {
  return(qplot(data = dataframe, x = 'data',
               y = column, geom = 'boxplot',
               xlab = '',
               ylab = ylab, color = I('black'), fill = I('#F79420')))
}
grid.arrange(boxplot_function(dataframe$fixed.acidity, 'fixed acidity'),
             boxplot_function(dataframe$volatile.acidity, 'volatile acidity'),
             boxplot_function(dataframe$citric.acid, 'citric acid'),
             boxplot_function(dataframe$Total.acidity, 'Total Acidity'),
             boxplot_function(dataframe$residual.sugar, 'residual sugar'),
             boxplot_function(dataframe$chlorides, 'chlorides'),
             boxplot_function(dataframe$free.sulfur.dioxide, 'free sulf. dioxide'),
             boxplot_function(dataframe$total.sulfur.dioxide, 'total sulf. dioxide'),
             boxplot_function(dataframe$density, 'density'),
             boxplot_function(dataframe$pH, 'pH'),
             boxplot_function(dataframe$sulphates, 'sulphates'),
             boxplot_function(dataframe$alcohol, 'alcohol'),
             ncol = 4)

```

####Boxplots for Two variable (Quality V/S Different Variables)
```{r,echo=FALSE}
# Creating Boxplot Function for Two Variables

two_variable_boxplot <- function(x, y, xlab ,ylab) {
  return(qplot(data = dataframe, x = x, y = y, 
               geom = 'boxplot', ylab = ylab, xlab=xlab,
               color = I('black'), fill = I('#F79420')))
}
grid.arrange(two_variable_boxplot(dataframe$quality, dataframe$fixed.acidity,'Quality','fixed acidity'),
             two_variable_boxplot(dataframe$quality, 
                                  dataframe$volatile.acidity,
                                   'Quality','volatile acidity'),
             two_variable_boxplot(dataframe$quality, 
                                  dataframe$citric.acid,
                                   'Quality','citric acid'),
             two_variable_boxplot(dataframe$quality, 
                                  dataframe$Total.acidity,
                                   'Quality','Total acidity'),
             two_variable_boxplot(dataframe$quality, y =             
                                    log10(dataframe$residual.sugar),
                                   'Quality','residual sugar'),
             two_variable_boxplot(dataframe$quality, 
                                  log10(dataframe$chlorides),
                                   'Quality','chlorides'),
             two_variable_boxplot(dataframe$quality,     
                                  dataframe$free.sulfur.dioxide,
                                   'Quality','free SO2'),
             two_variable_boxplot(dataframe$quality, 
                                  dataframe$total.sulfur.dioxide,
                                   'Quality','total SO2'),
             two_variable_boxplot(dataframe$quality, 
                                  dataframe$density,
                                   'Quality','density'),
             two_variable_boxplot(dataframe$quality, 
                                  dataframe$pH,
                                   'Quality','pH'),
             two_variable_boxplot(dataframe$quality, 
                                  log10(dataframe$sulphates),
                                   'Quality','sulphates'),
             two_variable_boxplot(dataframe$quality, 
                                  dataframe$alcohol,
                                   'Quality','alcohol'),
             ncol = 4)
```

####Boxplots for Two variable (Rating V/S Different Variables)

```{r,echo=FALSE}
grid.arrange(two_variable_boxplot(dataframe$rating, 
                                  dataframe$fixed.acidity,
                                   'Rating','fixed acidity'),
             two_variable_boxplot(dataframe$rating,
                                  dataframe$volatile.acidity,
                                    'Rating','volatile acidity'),
             two_variable_boxplot(dataframe$rating, 
                                  dataframe$citric.acid,
                                    'Rating','citric acid'),
             two_variable_boxplot(dataframe$rating, 
                                  dataframe$Total.acidity,
                                    'Rating','Total acidity'),
             two_variable_boxplot(dataframe$rating, log10(dataframe$residual.sugar),
                                    'Rating','residual sugar'),
             two_variable_boxplot(dataframe$rating, 
                                  log10(dataframe$chlorides),
                                    'Rating','chlorides'),
             two_variable_boxplot(dataframe$rating, 
                                  dataframe$free.sulfur.dioxide,
                                    'Rating','free SO2'),
             two_variable_boxplot(dataframe$rating, 
                                  dataframe$total.sulfur.dioxide,
                                    'Rating','total SO2'),
             two_variable_boxplot(dataframe$rating, 
                                  dataframe$density,
                                    'Rating','density'),
             two_variable_boxplot(dataframe$rating, 
                                  dataframe$pH,
                                    'Rating','pH'),
             two_variable_boxplot(dataframe$rating, 
                                  log10(dataframe$sulphates),
                                    'Rating','sulphates'),
             two_variable_boxplot(dataframe$rating, 
                                  dataframe$alcohol,
                                   'Rating','alcohol'),
             ncol = 4)

```

#### Analysis of Boxplots

By examining the plots, some conclusions can be drawn:

```{r,echo=FALSE}
two_variable_boxplot(dataframe$rating, 
                     dataframe$fixed.acidity,
                     'Rating','fixed acidity')
```

- A `good` wine causes because of high `fixed.acidity` (tartaric, malic, citric, and succinic acids).

```{r,echo=FALSE}
two_variable_boxplot(dataframe$rating, 
                     dataframe$citric.acid,
                    'Rating','citric acid')
```

- A `good` wine causes because of high `citric.acid`.

```{r,echo=FALSE}
two_variable_boxplot(dataframe$rating, 
                     dataframe$volatile.acidity,
                    'Rating','volatile acidity')
```

- A `good` wine causes because of low `volatile.acidity`(acetic acid).

```{r,echo=FALSE}
two_variable_boxplot(dataframe$rating, 
                     dataframe$pH,
                     'Rating','pH')
```

- A `good` wine causes because of lower `pH` value, means higher acidity.

```{r,echo=FALSE}
two_variable_boxplot(dataframe$rating, 
                     log10(dataframe$sulphates),
                     'Rating','sulphates')
```

- A `good` wine causes because of higher `sulphates`.

```{r,echo=FALSE}
two_variable_boxplot(dataframe$rating, 
                     log10(dataframe$sulphates),
                     'Rating','alcohal')
```

- A `good` wine causes because of higher `alcohal`.
- Little bit, lesser `chlorides` and `density` also affect the wine quality.
- `SO2` and `Residual Sugar` donot seems to have any affect on wine quality.

#### Correlation in between Quality and Different Variables

```{r,echo=FALSE}

#Function for Correlation Test

correlation_test <- function(x, y) {
  return(cor.test(x, as.numeric(y),method = 'pearson')$estimate)
}
  correlations <- c( 
  correlation_test(dataframe$fixed.acidity, dataframe$quality),
  correlation_test(dataframe$volatile.acidity, dataframe$quality),
  correlation_test(dataframe$citric.acid, dataframe$quality),
  correlation_test(dataframe$Total.acidity, dataframe$quality),
  correlation_test(log10(dataframe$residual.sugar), dataframe$quality),
  correlation_test(log10(dataframe$chlorides), dataframe$quality),
  correlation_test(dataframe$free.sulfur.dioxide, dataframe$quality),
  correlation_test(dataframe$total.sulfur.dioxide, dataframe$quality),
  correlation_test(dataframe$density, dataframe$quality),
  correlation_test(dataframe$pH, dataframe$quality),
  correlation_test(log10(dataframe$sulphates), dataframe$quality),
  correlation_test(dataframe$alcohol, dataframe$quality))
  
  names(correlations) <- c('fixed.acidity', 
                           'volatile.acidity', 
                           'citric.acid',
                           'Total.acidity', 
                           'log10.residual.sugar',
                           'log10.chlordies', 
                           'free.SO2',
                           'total.SO2', 
                           'density', 
                           'pH',
                           'log10.sulphates', 
                           'alcohol')
  correlations
```

Strong Correlation found in quality and 

- Alcohal
- Volatile Acidity
- Log10 of Sulphates
- Citric Acid

Definetely Alcohal, Volatile Acidity, Log10 of Sulphates and Citric acid are the main parameters that affecting the quality of wine.

#### Scatter Plot

To graphically check the correlation and the represent the relatationship graphically, it required to plot the Scatter plot. 

```{r fig.width= 12,fig.height=8,echo=FALSE}
  grid.arrange(ggplot(data = dataframe, aes(x = log10(sulphates), 
                               y = alcohol)) +
                              facet_wrap(~rating) +
                              geom_point(color = 'orange'),
                ggplot(data = dataframe, aes(x = volatile.acidity,
                               y = alcohol)) +
                              facet_wrap(~rating) +
                              geom_point(color = 'orange'),
                ggplot(data = dataframe, aes(x = citric.acid, 
                               y = alcohol)) +
                               facet_wrap(~rating) +
                               geom_point(color = 'orange'),
                ggplot(data = dataframe, aes(x = volatile.acidity, 
                               y = log10(sulphates))) +
                              facet_wrap(~rating) +
                               geom_point(color = 'orange'),
                ggplot(data = dataframe, aes(x = citric.acid, 
                               y = log10(sulphates))) +
                               facet_wrap(~rating) +
                               geom_point(color = 'orange'),
                ggplot(data = dataframe, aes(x = citric.acid, 
                               y = volatile.acidity)) +
                                  facet_wrap(~rating) +
                                  geom_point(color = 'orange'),
                ggplot(data = dataframe, aes(x = fixed.acidity, 
                               y = citric.acid)) +
                              geom_point(color = 'orange'),
                ggplot(data = dataframe, aes(x = volatile.acidity, 
                               y = citric.acid)) +
                               geom_point(color = 'orange'),
                ggplot(data = dataframe, aes(x = log10(Total.acidity), 
                               y = pH)) +
                               geom_point(color = 'orange'),
                 ggplot(data = dataframe, aes(x = free.sulfur.dioxide, 
                               y = total.sulfur.dioxide)) +
                               geom_point(color = 'orange'))
  
```

```{r echo=FALSE}
  correlation_scatter_plot <- c( 
  correlation_test(log10(dataframe$sulphates), dataframe$alcohol),
  correlation_test(dataframe$volatile.acidity, dataframe$alcohol),
  correlation_test(dataframe$volatile.acidity, log10(dataframe$sulphates)),
  correlation_test(dataframe$citric.acid, log10(dataframe$sulphates)),
  correlation_test(dataframe$citric.acid, dataframe$volatile.acidity),
  correlation_test(dataframe$fixed.acidity, dataframe$citric.acid),
  correlation_test(dataframe$volatile.acidity, dataframe$citric.acid),
  correlation_test(log10(dataframe$Total.acidity), dataframe$pH),
  correlation_test(log10(dataframe$free.sulfur.dioxide), 
           dataframe$total.sulfur.dioxide))
  
   names(correlation_scatter_plot ) <- c('log10(Sulphates) v/s Alcohol', 
                           'volatile.acidity v/s Alcohol', 
                           'volatile.acidity v/s log10(sulphates)',
                           'citric.acid v/s log10 (sulphates)', 
                           'citric.acid vs volatile.acidity',
                           'fixed.acidity v/s citric.acid',
                           'volatile.acidity v/s citric.acid',
                           'Total.acidity v/s pH',
                           'free SO2 v/s Total SO2')
  correlation_scatter_plot 
  
  
```  

I found very strong correlation between `Total.acidity` and `pH`, and the strongest correlation after that is between `free.SO2` and `total.SO2`. 

This is obvious that the strong correlation found in Acidity and pH because of pH is inversely proportional to acidity. One question need to ask what other components affecting pH other than acids. We can see this difference by building a linear model.

#### Errors in between Observerd pH and Predictive pH
```{r echo=TRUE,warning=FALSE}
#Defining Linear Model
m <- lm((pH) ~ log10(Total.acidity),data=dataframe)
#Making predictions according to LM.
dataframe$pH.predictions <- predict(m, dataframe)  
#Calculating % of Error
dataframe$pH.error <- ((dataframe$pH.predictions - dataframe$pH)/dataframe$pH)*100
#Creating a Boxplot in between Quality V/S pH.error
qplot(data = dataframe, 
      x = dataframe$quality, y = dataframe$pH.error, 
      geom = 'boxplot', ylab = 'pH Error(in %)', 
      xlab= 'Quality',color = I('black'),  fill = I('#F79420')) + 
  scale_y_continuous(limits = c(-10,10), breaks = seq(-10,10,1))

```  
  
The median % error in all cases remain somewhere close to 0 except in quality of 3. The median % of error of quality of 3 is at slightly lower than negative 4%. In this case we can conclude, the acids are not performing well in this case, or error is caused because of impurity in Acids.


# Summary and Plots

## Effect of acids on Wine

```{r echo=FALSE, Final_plot_one}

grid.arrange(ggplot(data = dataframe, aes(x = quality, 
                                          y = fixed.acidity,
                                   color = quality)) + 
               ylab('Fixed Acidity (g/dm^3)') +
               xlab('Quality') +
               geom_boxplot(),
             ggplot(data = dataframe, aes(x = quality, 
                                          y = volatile.acidity,
                                   color = quality)) +
               ylab('Volatile Acidity (g/dm^3)') +
               xlab('Quality') +
               geom_boxplot(), 
             ggplot(data = dataframe, aes(x = quality,
                                          y = citric.acid,
                                   color = quality)) +
               ylab('Citric Acid (g/dm^3)') +
               xlab('Quality') +
               geom_boxplot(), 
             ggplot(data = dataframe, aes(x = quality, 
                                          y = pH,
                                   color = quality)) +
               ylab('pH') +
               xlab('Quality') +
               geom_boxplot())
             
             grid.arrange(ggplot(data = dataframe, aes(x = rating, 
                                                       y = fixed.acidity,
                                   fill = rating)) + 
               ylab('Fixed Acidity (g/dm^3)') +
               xlab('Rating') +
               geom_boxplot(),
             ggplot(data = dataframe, aes(x = rating, 
                                          y = volatile.acidity,
                                   fill = rating)) +
               ylab('Volatile Acidity (g/dm^3)') +
               xlab('Rating') +
               geom_boxplot(), 
             ggplot(data = dataframe, aes(x = rating, 
                                          y = citric.acid,
                                   fill = rating)) +
               ylab('Citric Acid (g/dm^3)') +
               xlab('Rating') +
               geom_boxplot(), 
             ggplot(data = dataframe, aes(x = rating, 
                                          y = pH,
                                   fill = rating)) +
               ylab('pH') +
               xlab('Rating') +
               geom_boxplot())
```

#### Conclusions on effect of acidity on Wine

- There is little effect of Fixed acidity on Wine. But lower than 8 g/dm^3 acetic acid can cause a wine to be a `bad` wine, and a slight larger value than 8 g/dm^3 acetic acid can convert a wine into a `good` wine. 

- Volatile acidity which causes because of tartaric, malic, citric, and succinic acids should be slightly below 0.4 g/dm^3.

- If Citric acid is closer to 0 g/dm^3, then this can cause a wine to be `bad`. A value between 0.25-0.50 g/dm^3 is required for `good` wine.

- For a `good` wine pH value should be in between 3-3.5, but a closer value to 3.5 can cause a wine to be `bad`.

## Effects of Alcohol on Wine Quality

```{r echo=FALSE, Final_plot_two}
grid.arrange(ggplot(data = dataframe, aes(x = quality, 
                                          y = alcohol,
                      color = quality)) +
  geom_boxplot() +
  xlab('Quality') +
  ylab('Alcohol (% volume)'),ggplot(data = dataframe, aes(x = rating, 
                                                          y = alcohol,
                      fill = rating)) +
  geom_boxplot() +
  xlab('Rating') +
  ylab('Alcohol (% volume)'))
```  
  
The boxplot show a higher quantity of alcohol may cause a `good` quality wine. But, as the outliners show, alchohol content alone didnot produce high quality wine.

## What is the actual difference in Bad and Good Wine

To understand this, I removed `average` rating from plot.

```{r echo=FALSE, Plot_Three}
ggplot(data = subset(dataframe, rating != 'average'),
       aes(x = volatile.acidity, y = alcohol,
                      color = rating)) +
  geom_point() + 
    geom_line(x=0.45,linetype=2,color='orange')+
    xlab('Volatile Acidity (g / dm^3)') +
  ylab('Alcohol (% volume)')
```
  
  
- From this plot, this can be concluded the `good` wine creating a cluster when quantity of volatile.acidity is low and quantity of alcohal is high compare to `bad` wine (except some cases).
- The best way to make a `good` wine can be keeping volatile.acidity low and keeping alcohol high.


## Histogram Based on % Volume of Alcohol

```{r echo=TRUE, message=FALSE,rating_alcohol}
#Dividing data according to Alcohol %
dataframe$AlcoholRating <- ifelse(dataframe$alcohol < 9.5, 'Low', 
                           ifelse(dataframe$alcohol < 12, 'Moderate', 'High'))
#Adding levels
dataframe$AlcoholRating <- ordered(dataframe$AlcoholRating,
                     levels = c('Low', 'Moderate', 'High'))

summary(dataframe$AlcoholRating)

ggplot(data= dataframe, aes(x=as.factor(dataframe$quality) , fill = dataframe$AlcoholRating)) + 
  geom_histogram(aes(y=..count../sum(..count..)))  +
  xlab("Wine Quality") + ylab("Percentage") 
```


For plotting this, we divided the data into 3 section according to % of Alcohol according to this

- `Low` with less than 9.5%
- `Moderate` with inbetween 9.5 and 12%
- `High` above 12%.

I found as the quality increases, the % of alcohal also increases. The number of wine with `Low` % of alcohol vanished as Wine Quality increases.

#Summary

The findings for `Good` Wine as follows.

- A wine with greater 8 g/dm3 acetic acid (Fixed Acid) can cause a good wine.
- Volatile acid should be remain below 0.4 g/dm3.
- Citric Acid should be at 0.25-050 g/dm3.
- I dont able to find much difference of pH value in `bad` and `good` wine. But the value should be closer to 3.5
- Alcohol % should be above 9.5%


#Final Reflection

I was able to determine which factor affect the wine quality most i.e alcohol, sulphates and acidity. Important note is that quality or rating is subjective manner, and can change from person to person. The above graphs described what makes a wine `good`. 

I never done the statistical analysis from boxplot before, it took me time to understand the boxplot.

Furthur in most of the links, like this [Link](http://www.realsimple.com/holidays-entertaining/entertaining/food-drink/alcohol-content-wine), I found a Strong/High Wine is called when Alcohol % is called when wine is above 14.5%. But I found 0 data which is above 14.5%. I searched pretty hard then I found some 10 years ago, above 12% was considered as Strong Wine.

This is possible the dataset is old, or dataset is made from old parameters of wine.
I furthur faced many problems in many functions of R, as is grid.arrange, gridarrange and in creating functions. I pretty worked hard in understanding this in [stackoverflow](http://stackoverflow.com/).

On other note, the dataset is from which country is not defined. The various parameters defined differently in different country. Demographically and ethinically, a wine which is `bad` for one maybe `good` for other. For example, a French can rate a wine as `good` but a person from US maybe rate the same wine as `bad`.

##Future Analysis

- Future Analysis can be done by sub-classification is `residual.sugar` in `sweet wine` and `dry wine`. [Wikipedia](http://en.wikipedia.org/wiki/Sweetness_of_wine#Residual_sugar)
- Furthur examination of outliers.
- Rating and Quality based on region. 
